---
title: "Precipitation sinusoidal isocape model: data preparation"
author: "Andy McKenzie (NIWA), Bruce Dudley (NIWA)"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output:
  html_document: 
    theme: united
    code_folding: show
    toc: yes
    toc_float:
      collapsed: true   
    fig_caption: yes
    number_sections: yes
    df_print: kable
  word_document: 
    toc: yes
  pdf_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# Global options
#
# echo = TRUE. Show the code
# comment = "". Don't put anything in front of results from code
# warning = FALSE. Don't display warning messages
# cache = TRUE. Only process new code when you knit
# dev = 'png'. Save graphics in figure folder as png
# fig.height = 5. Height of figures (in inches)

knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  warning = FALSE,
  message = FALSE,
  cache = FALSE,
  dev = "png", dpi = 300,
  fig.height = 5
)

# tidyverse packages
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(lubridate)

# summary of data frame columns
library(skimr)

library(GGally) # for scatterplots with correlation with ggpairs()

# https://docs.stadiamaps.com/guides/migrating-from-stamen-map-tiles/#ggmap
# devtools::install_github("stadiamaps/ggmap”)
library(ggmap) # get_stamenmap() & ggmap
# register_stadiamaps("ff428357-71cc-4a08-8912-9111113265ec", write = TRUE)

library(ggrepel) # geom_text_repel()

# For map plotting
library(sf)

# devtools::install_github("ropensci/rnaturalearthhires")
library(rnaturalearthhires)
library(rnaturalearth)
library(rnaturalearthdata)

rm(list = ls())

source("auxiliary.code.R")

unlink("Output/Data", recursive = TRUE)
dir.create("Output/Data")
```

# Isotope data

## Raw data

From the directory:

`O:\FWWR1708\Working\Task 4 - Field\Isotope work\writing\Paper 2. Bruce lead\data\Bruce precip analysis`

The data file for measured data is "precip.iso.data_BD.csv".

For the measured data file, some characterisations and summaries follow.

```{r}
ddata <- readr::read_csv("data/precip.iso.data_BD.csv")
ddata.orig <- ddata

slice_head(ddata.orig, n = 5)
```

The variable column contains six "variables":
`r unique(ddata$variable)`.

```{r}
ddata |> count(variable)
```

The day of the month for the measurement isn't given, so the measurement is located in the middle of the month.  

```{r}
ddata <- filter(ddata.orig, variable == "d18O") |>
  rename(d18O = val, Site = site.number) |>
  mutate(Date = lubridate::make_date(year = year, month = month, day = 15))
```

d2H. 


```{r}
ddata2 <- filter(ddata.orig, variable == "dD") |>
  rename(dD = val, Site = site.number) |>
  mutate(Date = lubridate::make_date(year = year, month = month, day = 15))
```

```{r}
skimr::skim(ddata)
```
dD

```{r}
skimr::skim(ddata2)
```

## Locations of sites

Site information (e.g. lat, long, elevation, district) is in the file
"lookup_rainsite_VCSN.csv". Or at least the closest VCSN.

```{r data site information}
site.info.orig <- readr::read_csv("data/lookup_rainsite_VCSN.csv")

site.info <- site.info.orig |>
  dplyr::select(Site = site_numbe, area, lat, lon = long, height = ELEVATION)
slice_head(site.info, n = 5)
```

Site location is shown by North Island and South Island.

```{r fig.height=8}
iso_plot_site_locations(site.info, "Site", "lat", "lon", "NI")
iso_plot_site_locations(site.info, "Site", "lat", "lon", "SI")
```

## Plots of raw data

```{r raw d18O plotted, fig.height=10}
iso_plot_raw(ddata, "Site", "Date", "d18O",
  "Raw d18O by site (1-20)",
  sites = paste(1:20)
)
iso_plot_raw(ddata, "Site", "Date", "d18O",
  "Raw d18O by site (21-41)",
  sites = paste(21:41)
)
iso_plot_raw(ddata, "Site", "Date", "d18O",
  "Raw d18O by site (42-58)",
  sites = paste(42:58)
)
```



```{r raw dD plotted, fig.height=10}
iso_plot_raw(ddata2, "Site", "Date", "dD",
  "Raw dD by site (1-20)",
  sites = paste(1:20)
)
iso_plot_raw(ddata2, "Site", "Date", "dD",
  "Raw dD by site (21-41)",
  sites = paste(21:41)
)
iso_plot_raw(ddata2, "Site", "Date", "dD",
  "Raw dD by site (42-58)",
  sites = paste(42:58)
)
```

## Save final isotope data set

Sites with few data points are dropped, and NA measured values.

```{r}
ddata <- ddata |>
  filter(!(Site %in% c(8, 9, 12, 40, 45))) |>
  drop_na(d18O)
```


Same for dD.

```{r}
ddata2 <- ddata2 |>
  filter(!(Site %in% c(8, 9, 12, 40, 45))) |>
  drop_na(dD)
```


```{r}
save(ddata,  file = "ProcessedData/ddata.RData")
slice_head(ddata, n = 5)

save(ddata2, file = "ProcessedData/ddata2.RData")
slice_head(ddata2, n = 5)

```


# The VCSN climate and environmental data

## The virtual climate station network (VCSN) data

An overview of the virtual climate station data is available here:

<https://niwa.co.nz/climate/our-services/virtual-climate-stations>

"Virtual Climate station Network (VCSN) data are estimates of daily
rainfall, potential evapotranspiration, air and vapour pressure, maximum
and minimum air temperature, soil temperature, relative humidity, solar
radiation, wind speed and soil moisture on a regular (\~5km) grid
covering the whole of New Zealand. The estimates are produced every day,
based on the spatial interpolation of actual data observations made at
climate stations located around the country."

One form in which the VCSN data are available is a series of yearly
zipped up files in the directory.

"Q:/CLIMATE/vcsn_data/"

The CRS seems to be NZGD 1949

<https://one.niwa.co.nz/display/CLIDB/VCSN+Grid+Points>

As described here:

<https://www.linz.govt.nz/data/geodetic-system/datums-projections-and-heights/geodetic-datums/new-zealand-geodetic-datum-1949-nzgd1949>

With EPSG number of 27258

<https://epsg.io/27258>

## Extracting daily VCSN daily data

Some function were made to explore and extract the VCSN data:

(a) `vcsn_info()`
(b) `vcsn_agent_locations()`
(c) `vcsn_append_nearest_vcsn()`
(d) `vcsn_append_climate_information()`

These functions are used below.

Set up the directory with the VCSN climate data. Sometimes before
extracting data, the Q drive will need to be "woken up" by clicking on
it in Windows Explorer.

```{r}
# VCSN.directory <- "Q:/CLIMATE/vcsn_data/"
# vcsn_info(VCSN.directory)
```

Find all the vcsn climate station/agent locations. The output is an sf
class object, and can be plotted.

```{r plot vcsn agent locations}
# vcsn.agent.locations  <- vcsn_agent_locations(VCSN.directory) 
# save(vcsn.agent.locations, file = "ProcessedData/vcsn.agent.locations.RDdata")
load(file = "ProcessedData/vcsn.agent.locations.RDdata")
slice_head(vcsn.agent.locations, n = 5)
plot(vcsn.agent.locations["Agent"], cex = 0.5, pch = 16)
```


Append nearest agent information and daily climate data for 2007--2022.
This takes a while, so is done once and the results saved, then loaded.

```{r extract climate information for sites}
# years <- 2023
# site.info.plus.climate.data <- vcsn_append_climate_information(site.info,VCSN.directory,

#                                                       years)

# save(site.info.plus.climate.data, file = "output/site.info.plus.climate.data.RData")

load(file = "ProcessedData/site.info.plus.climate.data.RData")
slice_head(site.info.plus.climate.data, n = 5)
```

The climate fields are identified as

<https://one.niwa.co.nz/display/CLIDB/VCSN+Times+and+Field+Descriptions>

<https://one.niwa.co.nz/display/SYSTEMS/The+size+of+the+problem>

(a) MSLP (Mean Sea Level Pressure)
(b) PET (Potential Evapotranspiration)
(c) Rain (daily rainfall)
(d) RH (relative humidity)
(e) SoilM (soil moisture)
(f) ETmp (earth temperature at 10cm depth)
(g) Rad (solar radiation)
(h) TMax (maximum temperature)
(i) Tmin (minimum temperature)
(j) VP (vapour pressure)
(k) Wind (average wind speed at 10m above ground level)
(l) Rain_bc (Rainfall with bias correction)
(m) Tmax_N (maximum temperature via different method)
(n) Tmin_N (minimum temperature via different method)

Potential evaporation or potential evapotranspiration is defined as the
amount of evaporation that would occur if a sufficient water source were
available.

*Rain_bc*, *Tmax_n*, *Tmin_N* were introduced in 2019.


## Mean value from 2007-2022

Convert to long format and make a summary via the mean value 2007--2022 by site.

```{r convert climate information and summary over years}
site.info.plus.climate.data.long <- site.info.plus.climate.data |> 
  relocate(height, .after = "Date") |> 
  pivot_longer(height:Tmin_N, names_to = "Site quantity", values_to = "Value") |> 
  dplyr::select(-c("area", "lat", "lon"))

slice_head(site.info.plus.climate.data.long, n = 5)

# mean values over 2007-2022 (inclusive)
site.climate.summary <- site.info.plus.climate.data.long |> 
  group_by(Site, VCSN.lat, VCSN.lon, `Site quantity`) |> 
  summarise(mean = mean(Value, na.rm = TRUE)) |> 
  pivot_wider(names_from = `Site quantity`, values_from = mean) |> 
  ungroup()

slice_head(site.climate.summary, n = 5)

rm(site.info.plus.climate.data.long)
```

Calculate mean annual range of monthly temperatures:

Take monthly average daily maximum temperature for each site and month of the time series

Pick the highest month for each year

Pick the lowest for each year

Subtract lowest from highest for each year to give annual range

Take the average range across all years == mean annual range of monthly temperatures

```{r}
#  Convert to date if not already
site.info.plus.climate.data$Date <- as.Date(site.info.plus.climate.data$Date)
#  Get months
site.info.plus.climate.data$Month <- months(site.info.plus.climate.data$Date)
#  Get years
site.info.plus.climate.data$Year <- format(site.info.plus.climate.data$Date,format="%y")

#Get mean annual range of maximum daily temperatures (averaged for each month) each VCSN.Agent, month and year combination

aggy <- aggregate(TMax ~ VCSN.Agent + Site + Month + Year, data = site.info.plus.climate.data, mean)

aggyMAX <- aggregate(TMax ~ VCSN.Agent + Site + Year, data = aggy, max)
aggyMIN <- aggregate(TMax ~ VCSN.Agent + Site + Year, data = aggy, min)
aggyRANGE <- aggyMIN
aggyRANGE$AnnualTempRange <- aggyMAX$TMax - aggyMIN$TMax
annualranges <- aggregate(AnnualTempRange ~ VCSN.Agent + Site, data = aggyRANGE, mean)

site.climate.summary<-dplyr::left_join(site.climate.summary, annualranges, by = "Site")

rm(site.info.plus.climate.data)
rm(aggy)
rm(aggyMAX)
rm(aggyMIN)
rm(aggyRANGE)
```

Now make the same VCSN summary file for 2007-2022, but across all VCSN points so that we can use it to extrapolate regression relationships across the country



```{r}
slice_head(vcsn.agent.locations, n = 5) 
dim(vcsn.agent.locations)

#get list of all VCSN agent numbers and elevations

elevations<-read.csv("data/VCS_elevations.csv")
keeps<-subset(elevations, select = c(AGENT_NO, ELEVATION))
names(keeps)<-c("Agent", "elevation")
elevations<-keeps

# # test run
# #  vcsn.2007.2022 <- vcsn_combine_daily_data(VCSN.directory, years = 2007)
# # test run went fine but the line below took over a day and needs to be run on the modelling computer
# # vcsn.2007.2022 <- vcsn_combine_daily_data(VCSN.directory, years = 2007:2022)
# # save(vcsn.2007.2022, file = "output/vcsn.2007.2022.RData")

# load("output/vcsn.2007.2022.RData")

# # calculate annual ranges of monthly average temperatures

# #  Convert to date if not already
# vcsn.2007.2022$Date <- as.Date(vcsn.2007.2022$Date)
# #  Get months
# vcsn.2007.2022$Month <- months(vcsn.2007.2022$Date)
# #  Get years
# vcsn.2007.2022$Year <- format(vcsn.2007.2022$Date,format="%y")

#  # Get mean annual range of maximum daily temperatures (averaged for each month) for each VCSN.Agent, month and year combination

# aggy1 <- aggregate(TMax ~ Agent + Month + Year, data = vcsn.2007.2022, mean)
#
# aggyMAX <- aggregate(TMax ~ Agent + Year, data = aggy1, max)
# aggyMIN <- aggregate(TMax ~ Agent + Year, data = aggy1, min)
#
# aggyRANGE<-aggyMIN
# aggyRANGE$AnnualTempRange<-aggyMAX$TMax - aggyMIN$TMax
#
# annualranges<-aggregate(AnnualTempRange ~ Agent, data = aggyRANGE, mean)


# vcsn.mean <- vcsn.2007.2022 |>
#   dplyr::select(-Tmin, -TMax,
#                 -Rain)  |>
#   rename(lat = Lat, lon = Longt) |>
#   group_by(Agent, lat, lon) |>
#   summarise(across(.cols = MSLP:Rain_bc, .fns = ~mean(.x, na.rm = TRUE)))
#
# national.climate.summary<-dplyr::left_join(vcsn.mean, annualranges, by = "Agent")
# #add elevations
# national.climate.summary<-dplyr::left_join(national.climate.summary, elevations, by = "Agent")
#
# names(vcsn.mean)
# vcsn.mean<-national.climate.summary[,1:14]
# names(vcsn.mean)<-c("Agent", "lat", "lon", "MSLP", "PET" , "RH"      ,        "SoilM"    ,       "ETmp"
#  ,"Rad"     ,        "VP"      ,        "Wind"      ,      "Rain_bc"   ,      "AnnualTempRange", "height")


# removing large R objects
# rm(vcsn.2007.2022)
# rm(aggy1)
# rm(aggyMAX)
# rm(aggyMIN)
# rm(aggyRANGE)

# save(national.climate.summary, file = "ProcessedData/national.climate.summary.RData")
# save(vcsn.mean, file = "ProcessedData/vcsn.mean.RData")

load("ProcessedData/national.climate.summary.RData")
load("ProcessedData/vcsn.mean.RData")
```


Introduce an "island" column with values South Island/North Island. 

```{r}
site.climate.summary <- site.climate.summary |> 
  mutate(island = if_else(VCSN.lon <= 174.49 & VCSN.lat <= -40.18, 
                          "South Island",
                          "North Island")) |> 
  relocate(island, .after = Site) 
```

Change some names and add the same "island" column for the national VCSN dataframe for mapping later 

```{r}
names(national.climate.summary)<-c("VCSN.Agent", "VCSN.lat", "VCSN.lon", "MSLP", "PET", "RH", "SoilM", "ETmp", 
 "Rad", "VP", "Wind", "Rain_bc", "AnnualTempRange", "height")

national.climate.summary <- national.climate.summary |> 
  mutate(island = if_else(VCSN.lon <= 174.49 & VCSN.lat <= -40.18, 
                          "South Island",
                          "North Island")) |> 
  relocate(island, .after = height) 
```

Make climate summary by site data frame into an sf object. Make climate summary by site data frame into an sf object. 

```{r}
site.climate.summary.sf <- site.climate.summary |> 
  sf::st_as_sf(coords = c("VCSN.lon", "VCSN.lat"), crs = 4326)
```

Make national climate summary into an sf object...and take a look at the annual temperature range nationally

```{r}
vcsn.agent.locations1 <- vcsn.agent.locations
names(vcsn.agent.locations1) <- c("VCSN.Agent", "vcsn.lon", "vcsn.lat", "geometry")
national.climate.summary.sf <- dplyr::full_join(national.climate.summary, 
                                                vcsn.agent.locations1,
                                                by = "VCSN.Agent")
national.climate.summary.sf <- st_as_sf(national.climate.summary.sf)

plot(national.climate.summary.sf["AnnualTempRange"], cex = 0.5, pch = 16)
```



## Plots of mean rainfall and soil moisture at sites

Plot showing mean rainfall and soil moisture at sites. Both have 
potential to be proxies for orographic effects due to mountain ranges.

*Rain_bc* the bias corrected version of *Rain* is used. 

```{r plot simple of rainfall}
plot(site.climate.summary.sf["Rain_bc"], cex = 0.5, pch = 16)
```

```{r nzmap sf object1}
nzmap <- rnaturalearth::ne_countries(scale = 'large', 
                      country = "New Zealand",
                      returnclass = 'sf')
```


```{r plot rainfall with NZ map}
ggplot() + 
    geom_sf(data = nzmap, fill = "tan1") + 
    geom_sf(data = site.climate.summary.sf,
            mapping = aes(size = Rain_bc, colour = Rain)) +
    coord_sf(xlim = c(164.8, 179.4), ylim = c(-47.7, -33.8)) +
    scale_colour_viridis_c() +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("Mean daily rainfall (2007-2022)")
```


```{r plot soil moisture with NZ map}
ggplot() + 
    geom_sf(data = nzmap, fill = "tan1") + 
    geom_sf(data = site.climate.summary.sf,
            mapping = aes(size = SoilM, colour = SoilM)) +
    coord_sf(xlim = c(164.8, 179.4), ylim = c(-47.7, -33.8)) +
    scale_colour_viridis_c() +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("Mean soil moisture (2007-2022)")
```


```{r plot negative soil moisture with NZ map}
ggplot() + 
    geom_sf(data = nzmap, fill = "tan1") + 
    geom_sf(data = site.climate.summary.sf,
            mapping = aes(size = -SoilM, colour = -SoilM)) +
    coord_sf(xlim = c(164.8, 179.4), ylim = c(-47.7, -33.8)) +
    scale_colour_viridis_c() +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("Negative mean soil moisture (2007-2022)")
```

The soil moisture plateaus as rainfall increases. 

```{r plot mean soil moisture vs mean rainfall}
ggplot(site.climate.summary.sf, aes(x = Rain_bc, y = SoilM)) +
  geom_point(colour = "blue") +
  xlab("Mean daily rainfall") +
  ylab("Mean soil moisture") +
  ggtitle("Sites: soil moisture vs rainfall (mean 2007-09)")
```




## Correlation between VCSN predictor variables

There is high correlation between many VCSN variables, and for the purposes
of latter regression/correlation analyses with estimated sinusoidal parameters, 
some VCSN variables can be dropped. 

Concentrating on "temperature" variables, *Tmin* and *TMax* are highly
correlated with the variables *Tmin_N*, and *Tmax_n*, which in turn are
highly correlated with *ETmp* (earth temperature at 10cm depth).

So *ETmp* is retained but the other four are dropped (*Tmin*, *TMax*, *Tmin_N*, *Tmax_N*).

We will also retain the *AnnualTempRange* metric

```{r correlation between temperature site characteristics}
site.T <- site.climate.summary |> 
  dplyr::select(ETmp, Rad, TMax, Tmin, Tmax_N, Tmin_N)

GGally::ggpairs(site.T)
corrplot::corrplot(cor(site.T), diag = FALSE, order = "hclust" )
```

Concentrating on "non-temperature" variables. *Rain* is dropped and the
bias-corrected version *Rain_bc* is retained.

```{r correlation between non temperature site characteristics, cache=FALSE}
site.non.T <- site.climate.summary |> 
  dplyr::select(MSLP, PET, Rain, SoilM, VP, Wind, Rain_bc, height, AnnualTempRange)

GGally::ggpairs(site.non.T)
corrplot::corrplot(cor(site.non.T), diag = FALSE, order = "hclust" )
```

```{r correlation between remaining site characteristics, fig.height=10}
site.remain <- site.climate.summary |> 
  dplyr::select(-Site, -island, 
                -TMax, -Tmin, -Tmax_N, -Tmin_N,
                -Rain)

GGally::ggpairs(site.remain)
corrplot::corrplot(cor(site.remain), diag = FALSE, order = "hclust" )
```


# Saved data

These data are need for the next block of code (01-isoscape-model-fits-regressions.Rmd).

## d18O data


```{r}
save(ddata, file = "Output/Data/ddata.RData")
dim(ddata)
names(ddata)
```

## Site climate mean value data

Mean value of climate variable at a **site**, for the closest agent. Taking the 
mean value over all data for 2007 to 2022. 

```{r}
save(site.climate.summary, file = "Output/Data/site.climate.summary.RData")
dim(site.climate.summary)
names(site.climate.summary)
```

## Agent locations


```{r}
save(vcsn.agent.locations, file = "Output/Data/vcsn.agent.locations.RData")
dim(vcsn.agent.locations)
names(vcsn.agent.locations)
```

Same things, but slightly different column names.

```{r}
save(vcsn.agent.locations1, file = "Output/Data/vcsn.agent.locations1.RData")
dim(vcsn.agent.locations1)
names(vcsn.agent.locations1)
```


## Agent climate mean value data

Mean value of climate data at an **agent location**. Taking the mean over all data
from 2007 to 2022. Also includes height at the agent location.


```{r}
save(vcsn.mean, file = "Output/Data/vcsn.mean.RData")
dim(vcsn.mean)
names(vcsn.mean)
```

A slightly different version: 

(a) prefix of VCSN for "Agent","lat", "long",
(b) includes "island" column (South Island/North Island). 

```{r}
save(national.climate.summary, file = "Output/Data/national.climate.summary.RData")
dim(national.climate.summary)
names(national.climate.summary)
```





