---
title: "Precipitation sinusoidal isocape model: model fitting and regressions"
author: "Andy McKenzie (NIWA), Bruce Dudley (NIWA)"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output:
  html_document: 
    theme: united
    code_folding: show
    toc: yes
    toc_float:
      collapsed: true   
    fig_caption: yes
    number_sections: yes
    df_print: kable
  word_document: 
    toc: yes
  pdf_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# Global options
#
# echo = TRUE. Show the code
# comment = "". Don't put anything in front of results from code
# warning = FALSE. Don't display warning messages
# cache = TRUE. Only process new code when you knit
# dev = 'png'. Save graphics in figure folder as png
# fig.height = 5. Height of figures (in inches)

knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  warning = FALSE,
  message = FALSE,
  cache = FALSE,
  dev = "png", dpi = 300,
  fig.height = 5
)


# tidyverse packages
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(lubridate)
library(purrr)

# summary of data frame columns
library(skimr)

library(GGally) # for scatterplots with correlation with ggpairs()

# https://docs.stadiamaps.com/guides/migrating-from-stamen-map-tiles/#ggmap
# devtools::install_github("stadiamaps/ggmap”)
library(ggmap) # get_stamenmap() & ggmap
# register_stadiamaps("ff428357-71cc-4a08-8912-9111113265ec", write = TRUE)

# For map plotting
library(sf)
library(stars)

# devtools::install_github("ropensci/rnaturalearthhires")
library(rnaturalearthhires)
library(rnaturalearth)
library(rnaturalearthdata)

library(ggrepel) # geom_text_repel()
library(cowplot) # multiple plots for a figure

library(caret) # For LOOCV

library(skimr) # summary of data

library(robustbase)
library(modelr)
library(dotwhisker) # dot whisker plot for multiple regression results
library(broom) # glance(), tidy() for linear regressions

# model kriggin
library(gstat)

# For reading in global isotope prediction maps
library(raster)

library(flextable)
set_flextable_defaults(big.mark = "")


unlink("Output/ModelFitResults", recursive = TRUE)
dir.create("Output/ModelFitResults")


rm(list = ls())
source("auxiliary.code.R")

ft <- function(input.data, caption = "", num.rows = Inf) {
  ft <- input.data |>
    slice_head(n = num.rows) |>
    flextable() |>
    autofit() |>
    theme_zebra() |>
    set_caption(caption)
  ft
}
```


# Load in data



```{r}
# d18O and dD data
load("ProcessedData/ddata.RData")
load("ProcessedData/ddata2.RData")
slice_head(ddata, n = 5)

# site characteristics
load("Output/Data/site.climate.summary.RData")
glimpse(site.climate.summary)

# agent locations
load("Output/Data/vcsn.agent.locations1.RData")
slice_head(vcsn.agent.locations1, n = 5)

# mean value of VCSN climate data at agent locations
load("Output/Data/vcsn.mean.RData")
glimpse(vcsn.mean)

# Same as vcsn.mean, but (a) Prefix VCSN for "Agent", "lat", 
load("Output/Data/national.climate.summary.RData")
glimpse(national.climate.summary)

site.info.orig <- readr::read_csv("data/lookup_rainsite_VCSN.csv")
site.info <- site.info.orig |>
  dplyr::select(Site = site_numbe, area, lat, lon = long, height = ELEVATION)
slice_head(site.info, n = 5)
```


# Fitting sinusoidal curve to each site

## Fitting procedure

A sinusoidal curve is fitted to each site (Allen et al. 2018, p. 4861)
of the form:

$$quantity = Amplitude*sin(2\pi*Julian.Day.Fraction - phase) + offset$$
where `Julian.Day.Fraction` is zero at the start of the year, and one at
the end.

If the phase is positive the sin curve is shifted to the right, and if
the phase is negative the sin curve is shifted to the left.

The fitting procedure details are:

(a) Non-linear fit with estimation of amplitude, phase, and offset
    (including estimation error)

(b) Emulating a previous robust fitting procedure, a **Huber** weighting
    is given to residuals. This can easily be changed to **bisquare**

(c) Fitted using the **nlrob** function from the *robustbase* R package.

(d) Bounds are set for the estimated parameters
    ($0 \le Amplitude \le 10$, $-\pi \le phase \le \pi$,
    $-20 \le offset \le 0$). If bounds are not set then for some sites a
    negative value is estimated for the amplitude, and large values for
    the phase.

The fitting function `iso_fit_sinusoidal_model()` needs to be supplied
the arguments:

1.  The data (ddata)
2.  The name of the column containing the site ID ("Site")
3.  The name of the column containing the date ("Date")
4.  The name of the column containing the quantity that is to be fitted
    to ("d18O")

```{r}
mods.d18O <- iso_fit_sinusoidal_model(ddata, "Site", "Date", "d18O",
  lower.limits = c(Amplitude = 0.1, phase = -3.1416, offset = -10),
  upper.limits = c(Amplitude = 5.0, phase = 3.1416, offset = -3)
)
```

Same for dD EXCEPT changed bounds to scale for the difference between dD and d18O. 
In NZ (and globally) dD = 8*d18O+10 so bounds changed to:

foe dD, bounds are set for the estimated parameters
    ($0 \le Amplitude \le 90$, $-\pi \le phase \le \pi$,
    $-150 \le offset \le -10$). If bounds are not set then for some sites a
    negative value is estimated for the amplitude, and large values for
    the phase.
    
```{r}
mods.dD <- iso_fit_sinusoidal_model(ddata2, "Site", "Date", "dD",
  lower.limits = c(Amplitude = 0.1, phase = -3.1416, offset = -150),
  upper.limits = c(Amplitude = 90, phase = 3.1416, offset = -10)
)
```

The output is a model object containing fitted parameters, residuals,
fitted sinusoidal curves, etc.

## Fitted curves

Fitted sinusoidal curves are shown for all sites, or selected sites.

```{r, fig.height=8}
iso_plot_fitted_curve(mods.dD, sites = paste(1:20))
iso_plot_fitted_curve(mods.dD, sites = paste(21:41))
iso_plot_fitted_curve(mods.dD, sites = paste(42:58))
```
For dD
```{r, fig.height=8}
iso_plot_fitted_curve(mods.dD, sites = paste(1:20))
iso_plot_fitted_curve(mods.dD, sites = paste(21:41))
iso_plot_fitted_curve(mods.dD, sites = paste(42:58))
```

## Residuals

Residuals are shown for each site, and all sites combined.

```{r, fig.height=12}
iso_plot_residuals(mods.d18O)
```
dD residuals for each site, and all sites combined.

```{r, fig.height=12}
iso_plot_residuals(mods.dD)
```

```{r}
iso_plot_residuals(mods.d18O, combined = TRUE)
```

## RMSE and MAD

Room mean square error (RMSE) and median absolute deviation (MAD).
Latter these are plotted on a map by site.

```{r}
RMSE.by.site <- iso_extract_RMSE_by_site(mods.d18O)
slice_head(RMSE.by.site, n = 5)
```
for dD

```{r}
RMSE.by.site.dD <- iso_extract_RMSE_by_site(mods.dD)
slice_head(RMSE.by.site.dD, n = 5)
```


```{r}
MAD.by.site <- iso_extract_MAD_by_site(mods.d18O)
slice_head(MAD.by.site, n = 5)
```

for dD

```{r}
MAD.by.site.dD <- iso_extract_MAD_by_site(mods.dD)
slice_head(MAD.by.site.dD, n = 5)
```

Plots below not repeated for dD

```{r}
plot.RMSE <- ggplot(data = RMSE.by.site, aes(x = Site, y = RMSE)) +
  geom_col(fill = "blue") +
  ggtitle("Root mean square error (RMSE) by site")
plot.RMSE
```

```{r}
plot.MAD <- ggplot(data = MAD.by.site, aes(x = Site, y = MAD)) +
  geom_col(fill = "blue") +
  ggtitle("Median absolute deviation/residual (MAD) by site")
plot.MAD
```

Interactive version of plots

```{r}
plotly::ggplotly(plot.RMSE, width = 800, height = 600)
```

```{r}
plotly::ggplotly(plot.MAD, width = 800, height = 600)
```

```{r}
plot.MAD <- ggplot(data = MAD.by.site, aes(x = Site, y = MAD)) +
  geom_col(fill = "blue") +
  ggtitle("Median absolute deviation/residual (MAD) by site")
plot.MAD
```

Interactive version of plot

```{r}
plotly::ggplotly(plot.MAD, width = 800, height = 600)
```

## Parameter estimates

### Tables

Parameter estimates with standard error, test-statistic (t-value), and
associated p-value.

```{r}
params.d18O.full <- iso_extract_parameters(mods.d18O)

params.d18O.full |>
  slice_head(n = 10) |>
  flextable() |>
  set_formatter(p.value = function(x) {
    formatC(x, format = "e", digits = 3)
  }) |>
  autofit() |>
  theme_zebra() |>
  set_caption("Parameter estimates and error (first ten rows).")

readr::write_csv(params.d18O.full, "Output/ModelFitResults/params.d18O.full.csv")
```

Or just the parameter estimates in a nicer looking format.

```{r}
params.d18O.est.only <- iso_extract_parameters(mods.d18O, estimates.only = TRUE)
ft(params.d18O.est.only, caption = "Parameter estimates only (first ten rows).", num.rows = 10)

readr::write_csv(params.d18O.est.only, "Output/ModelFitResults/params.d18O.est.only.csv")
```
same again for dD
```{r}
params.dD.full <- iso_extract_parameters(mods.dD)

params.dD.full |>
  slice_head(n = 10) |>
  flextable() |>
  set_formatter(p.value = function(x) {
    formatC(x, format = "e", digits = 3)
  }) |>
  autofit() |>
  theme_zebra() |>
  set_caption("Parameter estimates and error (first ten rows).")

readr::write_csv(params.dD.full, "Output/ModelFitResults/dDfits.csv")
```

Or just the parameter estimates in a nicer looking format.

```{r}
params.dD.est.only <- iso_extract_parameters(mods.dD, estimates.only = TRUE)
ft(params.dD.est.only, caption = "Parameter estimates only (first ten rows).", num.rows = 10)

readr::write_csv(params.dD.est.only, "Output/ModelFitResults/dDfits.est.only.csv")
```

### Plots

Whiskers plotted with 95% confidence intervals.

```{r , fig.height=8}
iso_plot_fitted_parameters(mods.d18O, "Amplitude")
```

```{r , fig.height=8}
iso_plot_fitted_parameters(mods.d18O, "phase")
```

```{r , fig.height=8}
iso_plot_fitted_parameters(mods.d18O, "offset")
```

# Correlations

## Between estimated parameters

There is some correlation between the estimated phase and offset
parameters. As the estimated phase value increases, the estimated offset
value decreases.

```{r data parameter estimates}
params.estimates <- iso_extract_parameters(mods.d18O, estimates.only = TRUE)

params.estimates |>
  dplyr::select(-Site) |>
  GGally::ggpairs()
```

## Between estimated parameters and basic site characteristics

```{r data combined parameters and site info}
# neg.phase.days is how many days the sin curve is shifted to the left
combine.data <- params.estimates |>
  left_join(site.info) |>
  mutate(neg.phase = -phase, .after = phase) |>
  mutate(neg.phase.days = -phase / (2 * pi) * 365.25, .after = neg.phase) |>
  mutate(neg.offset = -offset, .after = offset)

combine.data <- combine.data |>
  mutate(island = if_else(lon <= 174.49 & lat <= -40.18,
    "South Island",
    "North Island"
  ))

slice_head(combine.data, n = 5)
save(combine.data, file = "Output/Data/combine.data.RData")
```



```{r}
combine.vcsn <- combine.data |> 
  left_join(site.climate.summary)
```

```{r}
combine.vcsn.reg <- combine.vcsn |> 
  dplyr::select(-neg.phase, -neg.phase.days, -neg.offset, -lat, -lon, 
                -Tmin, -TMax, 
                -Tmin_N, -Tmax_N,
                -Rain)

slice_head(combine.vcsn.reg, n = 5)
save(combine.vcsn.reg, file = "Output/Data/combine.vcsn.reg.RData")
```





The *Amplitude* increases with longitude.

The *phase* decrease with latitude

The *offset* increases with latitude and longitude.

```{r plot combined correlation}
combine.data |>
  dplyr::select(-Site, -area) |>
  dplyr::select(-contains("neg")) |>
  GGally::ggpairs()
```

Shown below are some of the higher correlations of estimated parameters
with longitude, latitude, and height.

```{r plot amplitude vs lon}
ggplot(combine.data, aes(x = lon, y = Amplitude)) +
  geom_point(colour = "blue", size = 3) +
  xlab("Longitude") +
  ggtitle("Amplitude versus longitude")
```

```{r plot phase vs latitude}
ggplot(combine.data, aes(x = lat, y = phase)) +
  geom_point(colour = "blue", size = 3) +
  xlab("Latitude") +
  ggtitle("phase versus latitude")
```

```{r plot offset vs lat}
ggplot(combine.data, aes(x = lat, y = offset)) +
  geom_point(colour = "blue", size = 3) +
  xlab("Latitude") +
  ggtitle("offset versus latitude")
```

```{r plot offset vs height}
ggplot(combine.data, aes(x = height, y = offset)) +
  geom_point(colour = "blue", size = 3) +
  xlab("Height (m)") +
  ggtitle("offset versus height")
```


# Spatial distribution of estimated parameters

## Static maps

The area of circles is proportional to the quantity value, with the
colour shading binned.

The absolute value of the phase and offset are plotted (i.e. the
negative of their actual values), so the area is circle is proportional
to the absolute value of them.

For the phase, this is also converted to days by multiplying by 365.25.

A phase value of zero implies a peak for the sin curve one quarter of
the way through the year (about 91 days). All phase values are negative,
implying a shift of the sinusoidal curve to the left.

```{r}
s1 <- iso_plot_spatial_static(combine.data, "Site", "lat", "lon", "Amplitude")
s2 <- iso_plot_spatial_static(combine.data, "Site", "lat", "lon", "neg.phase")
s2.day <- iso_plot_spatial_static(combine.data, "Site", "lat", "lon", "neg.phase.days")
s3 <- iso_plot_spatial_static(combine.data, "Site", "lat", "lon", "neg.offset")

s1
s2
s2.day
s3

cowplot::plot_grid(s1, s3)
cowplot::plot_grid(s1, s2)
```

```{r}
```

Root mean square error by site

```{r}
RMSE.by.site.with.lat.lon <- iso_extract_RMSE_by_site(mods.d18O) |>
  left_join(site.info)
iso_plot_spatial_static(RMSE.by.site.with.lat.lon, "Site", "lat", "lon", "RMSE")
```

Median absolute deviation (i.e. residual) by site.

```{r}
MAD.by.site.with.lat.lon <- iso_extract_MAD_by_site(mods.d18O) |>
  left_join(site.info)
plot.MAD.by.site <- iso_plot_spatial_static(
  MAD.by.site.with.lat.lon,
  "Site", "lat", "lon", "MAD"
)
plot.MAD.by.site
```



# Correlation of sinusoidal predictors with VCSN predictors



## Amplitude

```{r}
iso_plot_spatial_static(combine.data, "Site", "lat", "lon", "Amplitude")
```

```{r plot amplitude vs VCSN predictors}
amp.long <- combine.vcsn.reg |> 
  dplyr::select(Amplitude, height, VCSN.lat: Wind, AnnualTempRange) |> 
  pivot_longer(cols = -Amplitude, names_to = "Predictor", values_to = "Value")

ggplot(amp.long, aes(x = Value, y = Amplitude)) +
  geom_point(colour = "blue") +
  facet_wrap(vars(Predictor), scale = "free_x")
```


```{r correlation Amplitude with Rad SoilM and annual temperature range, cache=FALSE}
corr.Amp.Rad <-   round(cor(combine.vcsn.reg$Amplitude, combine.vcsn.reg$Rad), 2)
corr.Amp.SoilM <- round(cor(combine.vcsn.reg$Amplitude, combine.vcsn.reg$SoilM), 2)
corr.Amp.AnnualTempRange <- round(cor(combine.vcsn.reg$Amplitude, combine.vcsn.reg$AnnualTempRange), 2)
```


The estimated sinusoidal Amplitude has highest correlation with
Rad (`r corr.Amp.Rad`) and SoilM (`r corr.Amp.SoilM`).

```{r}
combine.vcsn.reg |> 
  dplyr::select(Amplitude, height, ETmp: Wind, AnnualTempRange) |> 
  GGally::ggpairs()
```

Ignoring lat/lon a tree regression picks out SoilM as of importance for explaining 
variation in the Amplitude, following by interactions with MSLP and Rad. 

Note from Bruce: However, with AnnualTempRange added, the revised regression tree has Rad and AnnualTempRange as the top two predictors

```{r tree regression for amplitude}
tree.model <- tree::tree(Amplitude ~ height +
                 ETmp + MSLP + PET + Rad + 
                 Rain_bc + 
                 RH + SoilM +
                 VP + Wind + AnnualTempRange, 
                 data = combine.vcsn.reg)
plot(tree.model)
text(tree.model)
```

## Phase

```{r}
iso_plot_spatial_static(combine.data, "Site", "lat", "lon", "phase")
```

```{r plot phase versus VCSN }
phase.long <- combine.vcsn.reg |> 
  dplyr::select(phase, height, VCSN.lat: Wind, AnnualTempRange) |> 
  pivot_longer(cols = -phase, names_to = "Predictor", values_to = "Value")

ggplot(phase.long, aes(x = Value, y = phase)) +
  geom_point(colour = "blue") +
  facet_wrap(vars(Predictor), scale = "free_x")
```



```{r correlation phase with Rad and SoilM}
corr.phase.MSLP <-   round(cor(combine.vcsn.reg$phase, combine.vcsn.reg$MSLP), 2)
corr.phase.MSLP <- round(cor(combine.vcsn.reg$Amplitude, combine.vcsn.reg$VP), 2)
```


The estimated sinusoidal phase has highest correlation with the Mean Sea Level Pressure MSLP (`r corr.phase.MSLP`) and Vapour Pressure VP (`r corr.phase.MSLP`).

```{r plot correlation phase vs VCSN}
combine.vcsn.reg |> 
  dplyr::select(phase, height, ETmp: Wind, AnnualTempRange) |> 
  GGally::ggpairs()
```

VP and height are the two top VCSN predictors for phase

```{r tree regression for phase}
tree.model <- tree::tree(phase ~ height +
                 ETmp + MSLP + PET + Rad + 
                 Rain_bc + 
                 RH + SoilM +
                 VP + Wind + AnnualTempRange, 
                 data = combine.vcsn.reg)
plot(tree.model)
text(tree.model)
```

## Offset

```{r}
iso_plot_spatial_static(combine.data, "Site", "lat", "lon", "offset")
```

```{r}
offset.long <- combine.vcsn.reg |> 
  dplyr::select(offset, height, VCSN.lat: Wind, AnnualTempRange) |> 
  pivot_longer(cols = -offset, names_to = "Predictor", values_to = "Value")

ggplot(offset.long, aes(x = Value, y = offset)) +
  geom_point(colour = "blue") +
  facet_wrap(vars(Predictor), scale = "free_x")
```

VP, AnnualTempRange, ETmp and SoilM are the best predictors for offset
```{r tree regression for offset}
tree.model <- tree::tree(offset ~ height +
                 factor(island) +
                 VCSN.lat + VCSN.lon + 
                 ETmp + MSLP + PET + Rad + 
                 Rain_bc + 
                 RH + SoilM +
                 VP + Wind + AnnualTempRange, 
                 data = combine.vcsn.reg)
plot(tree.model)
text(tree.model)
```



# Second batch of regression models using VCSN data


## First - bring in Leave-one-out cross validation as a test of model suitability

```{r}

ctrl <- caret::trainControl(method = "LOOCV")
```


## Amplitude 

First try a basic model for amplitude, with and without height (elevation) as a predictor.
In this case, the model not including height has a slightly lower RMSE (and slightly higher r squared) than the one with height. 

```{r}
model <- train(Amplitude ~ VCSN.lat + VCSN.lon + height, 
               data = combine.vcsn.reg, 
               method = "lm", 
               trControl = ctrl)
print(model)

model2 <- train(Amplitude ~ VCSN.lat + VCSN.lon, 
                data = combine.vcsn.reg, 
                method = "lm",
                trControl = ctrl)
print(model2)

```


At this point we tried a range of different models. We used stepAIC to trim down models, then assessed the selected models based on LOOCV. For each of amplitude, phase, and offset we keep the model matching Allen (2018), and one other model (generally with more predictors). For brevity I have deleted the ones we chose not to keep. 

Below we test the model matching that used by Allen et al. (2018) for amplitude. 

Again, we use LOOCV to test the full model against the reduced one selected by stepAIC. 

```{r}
model3 <- train(Amplitude ~ VCSN.lat + VCSN.lon + height + Rain_bc + AnnualTempRange,
                data = combine.vcsn.reg,
                method = "lm",
                trControl = ctrl)
print(model3)

model4<- train(Amplitude ~ VCSN.lon + height + AnnualTempRange, 
               data = combine.vcsn.reg, 
               method = "lm", 
               trControl = ctrl)
print(model4)

```

This is the Allen et al. method model.

```{r}
Amp.mod1.2 <- lm(Amplitude ~ VCSN.lat + VCSN.lon + height + Rain_bc + AnnualTempRange, 
                 data = combine.vcsn.reg)

Amp.mod1.2step <- MASS::stepAIC(Amp.mod1.2)
summary(Amp.mod1.2step)
```


Here (for amplitude) is the 'full' model with all the less-correlated VCSN predictors plus height, lat, long and AnnualTempRange. The best model based on LOOCV is the STEPAIC-reduced model that comes from these predictors: model 6 - or A.mod1.step

```{r}
model5 <- train(Amplitude ~ height +
                 VCSN.lat + VCSN.lon + VCSN.lat:VCSN.lon +
                 ETmp + MSLP + PET + Rad + 
                 Rain_bc + 
                 RH + SoilM +
                 VP + Wind + AnnualTempRange, 
                data = combine.vcsn.reg, 
                method = "lm", 
                trControl = ctrl)
print(model5)

model6<- train(Amplitude ~ height + VCSN.lat + VCSN.lon + PET + Rad + VP + Wind, 
    data = combine.vcsn.reg,
    method = "lm", 
    trControl = ctrl)
print(model6)
```


```{r}
A.mod1 <- lm(Amplitude ~  height +
                 VCSN.lat + VCSN.lon + VCSN.lat:VCSN.lon +
                 ETmp + MSLP + PET + Rad + 
                 Rain_bc + 
                 RH + SoilM +
                 VP + Wind + AnnualTempRange,
                 data = combine.vcsn.reg)

A.mod1.step <- MASS::stepAIC(A.mod1, direction = "both", trace = FALSE)
summary(A.mod1.step)
car::vif(A.mod1.step)
car::Anova(A.mod1.step)
```


## Phase

Model matching that used by Allen et al. (2018) now applied to phase.

```{r}
model7 <- train(phase ~ VCSN.lat + VCSN.lon + height + Rain_bc + AnnualTempRange,
                data = combine.vcsn.reg, 
                method = "lm", 
                trControl = ctrl)
print(model7)

model8<- train(phase ~ VCSN.lat + VCSN.lon, 
               data = combine.vcsn.reg, 
               method = "lm", 
               trControl = ctrl)
print(model8)
```


```{r}
phase.mod1.2 <- lm(phase ~ VCSN.lat + VCSN.lon + height + Rain_bc + AnnualTempRange,
                   data = combine.vcsn.reg)

phase.mod1.2step <- MASS::stepAIC(phase.mod1.2)
summary(phase.mod1.2step)
```

And again for the 'full' phase model. Model 9 includes all predictors and model 10 gives the StepAIC-reduced model.

```{r}
model9 <- train(phase ~  height +
                 island +
                 VCSN.lat + VCSN.lon + VCSN.lat:VCSN.lon +
                 ETmp + MSLP + PET + Rad + 
                 Rain_bc + 
                 RH + SoilM +
                 VP + Wind + AnnualTempRange, 
                data = combine.vcsn.reg, 
                method = "lm", trControl = ctrl)
print(model9)

model10<- train(phase ~ height +
                 island +
                 VCSN.lat + VCSN.lon + VCSN.lat:VCSN.lon +
                 MSLP + Rad + SoilM +
                 VP + Wind + AnnualTempRange, 
                 data = combine.vcsn.reg, 
                 method = "lm", 
                 trControl = ctrl)
print(model10)
```

```{r}
phase.mod1 <- lm(phase ~  height +
                 island +
                 VCSN.lat + VCSN.lon + VCSN.lat:VCSN.lon +
                 ETmp + MSLP + PET + Rad + 
                 Rain_bc + 
                 RH + SoilM +
                 VP + Wind + AnnualTempRange,
                 data = combine.vcsn.reg)

phase.mod1.step <- MASS::stepAIC(phase.mod1, direction = "both", trace = FALSE)
summary(phase.mod1.step)
car::vif(phase.mod1.step, type = "predictor")
car::Anova(phase.mod1.step)
```

##Offset


Model matching that used by Allen et al. (2018) now applied to offset. First test both full and reduced models using LOOCV. 

And again for the 'full' phase model. Model 9 includes all predictors and model 10 gives the StepAIC-reduced model

```{r}
model11 <- train(offset ~ VCSN.lat + VCSN.lon + height + Rain_bc + AnnualTempRange,
                 data = combine.vcsn.reg,
                 method = "lm", 
                 trControl = ctrl)
print(model11)

model12<- train(offset ~ VCSN.lat + height + Rain_bc + AnnualTempRange, 
                data = combine.vcsn.reg, 
                method = "lm", 
                trControl = ctrl)
print(model12)
```

In the case below for the Allen predictors, the full model gave a slightly better LOOCV result. We'll keep that one. 

```{r}
offset.mod1.2 <- lm(offset ~ VCSN.lat + VCSN.lon + height + Rain_bc + AnnualTempRange, 
                    data = combine.vcsn.reg)

offset.mod1.2step <- MASS::stepAIC(offset.mod1.2)
summary(offset.mod1.2step)
```

And finally the 'full' model for offset. Again the StepAIC model gives the best LOOCV result. This StepAIC model doesn't include latitude, but does include Etmp as a predictor. 

```{r}
model13 <- train(offset ~  height +
                 island +
                 VCSN.lat + VCSN.lon + VCSN.lat:VCSN.lon +
                 ETmp + MSLP + PET + Rad + 
                 Rain_bc + 
                 RH + SoilM +
                 VP + Wind + AnnualTempRange, data = combine.vcsn.reg, method = "lm", trControl = ctrl)
print(model13)

model14<- train(offset ~  height +
                 island +
                 VCSN.lon +
                 ETmp + MSLP +
                 Rain_bc + 
                 RH +
                 VP + AnnualTempRange, 
                 data = combine.vcsn.reg, 
                 method = "lm", 
                 trControl = ctrl)
print(model14)

```


Here are the models for later use. 

```{r}
offset.mod1 <- lm(offset ~  height +
                 island +
                 VCSN.lat + VCSN.lon + VCSN.lat:VCSN.lon +
                 ETmp + MSLP + PET + Rad + 
                 Rain_bc + 
                 RH + SoilM +
                 VP + Wind + AnnualTempRange,
               data = combine.vcsn.reg)

offset.mod1.step <- MASS::stepAIC(offset.mod1, direction = "both", trace = FALSE)
summary(offset.mod1.step)
car::vif(offset.mod1.step, type = "predictor")
car::Anova(offset.mod1.step)
```



# Prediction over New Zealand 1 - linear model predictions

Apply the StepAIC-selected 'full' linear models made using the measurement site results across the VCSN grid. 

First amplitude. Also, set minimum amplitude limit to zero

```{r}
national.climate.summary$amp.pred <- predict(A.mod1.step, 
                                             newdata = national.climate.summary)

range(national.climate.summary$amp.pred)
national.climate.summary$amp.pred <- pmax(national.climate.summary$amp.pred, 0)
range(national.climate.summary$amp.pred)
```


Now phase and offset

```{r}

national.climate.summary$phase.pred<-predict(phase.mod1.step, 
                                             newdata = national.climate.summary)

range(national.climate.summary$phase.pred)
national.climate.summary$offset.pred<-predict(offset.mod1.step, newdata = national.climate.summary)
range(national.climate.summary$offset.pred)

```

Now all of the regressions following Allen et. al. (2018) for comparison later

```{r}
national.climate.summary$allen.amp.pred<-predict(Amp.mod1.2step, newdata = national.climate.summary)
range(national.climate.summary$allen.amp.pred)

national.climate.summary$allen.phase.pred<-predict(phase.mod1.2step, newdata = national.climate.summary)
range(national.climate.summary$allen.phase.pred)

national.climate.summary$allen.offset.pred<-predict(offset.mod1.2, newdata = national.climate.summary)
range(national.climate.summary$allen.offset.pred)
```


# Plots of some linear model predictions 



```{r}
national.climate.summary.sf <- dplyr::full_join(national.climate.summary, 
                                                vcsn.agent.locations1, 
                                                by = "VCSN.Agent") |>
                                sf::st_as_sf()

plot(national.climate.summary.sf["amp.pred"], cex = 0.5, pch = 16)
plot(national.climate.summary.sf["allen.amp.pred"], cex = 0.5, pch = 16)
plot(national.climate.summary.sf["offset.pred"], cex = 0.5, pch = 16)

```
# Calculation of residuals for 6 linear models



First, apply the national linear model predictions to the measurement sites only. 

```{r}
combine.vcsn.reg$amp.pred<-predict(A.mod1.step, newdata = combine.vcsn.reg)
combine.vcsn.reg$phase.pred<-predict(phase.mod1.step, newdata = combine.vcsn.reg)
combine.vcsn.reg$offset.pred<-predict(offset.mod1.step, newdata = combine.vcsn.reg)

combine.vcsn.reg$allen.amp.pred<-predict(Amp.mod1.2step, newdata = combine.vcsn.reg)
combine.vcsn.reg$allen.phase.pred<-predict(phase.mod1.2step, newdata = combine.vcsn.reg)
combine.vcsn.reg$allen.offset.pred<-predict(offset.mod1.2step, newdata = combine.vcsn.reg)
```


Then, calculate residuals for the various 'complete' models 

```{r}
combine.vcsn.reg$amp.pred.err<- combine.vcsn.reg$amp.pred - combine.vcsn.reg$Amplitude
combine.vcsn.reg$phase.pred.err<- combine.vcsn.reg$phase.pred - combine.vcsn.reg$phase
combine.vcsn.reg$offset.pred.err<- combine.vcsn.reg$offset.pred - combine.vcsn.reg$offset

combine.vcsn.reg$allen.amp.pred.err<- combine.vcsn.reg$allen.amp.pred - combine.vcsn.reg$Amplitude
combine.vcsn.reg$allen.phase.pred.err<- combine.vcsn.reg$allen.phase.pred - combine.vcsn.reg$phase
combine.vcsn.reg$allen.offset.pred.err<- combine.vcsn.reg$allen.offset.pred - combine.vcsn.reg$offset
```


#Mapping residuals for linear model predictions

Make regression summary by site data frame into an sf object. 

```{r}
site.regression.summary.sf <- combine.vcsn.reg |> 
  sf::st_as_sf(coords = c("VCSN.lon", "VCSN.lat"), crs = 4326)
```

```{r nzmap sf object}
nzmap <- rnaturalearth::ne_countries(scale = 'large', 
                      country = "New Zealand",
                      returnclass = 'sf')
```


```{r plot amplitude residuals with NZ map}
ggplot() + 
    geom_sf(data = nzmap, fill = "tan1") + 
    geom_sf(data = site.regression.summary.sf,
            mapping = aes(size = amp.pred.err, colour = Amplitude)) +
    coord_sf(xlim = c(164.8, 179.4), ylim = c(-47.7, -33.8)) +
    scale_colour_viridis_c() +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("Linear model amplitude residuals")

ggplot() + 
    geom_sf(data = nzmap, fill = "tan1") + 
    geom_sf(data = site.regression.summary.sf,
            mapping = aes(size = allen.amp.pred.err, colour = Amplitude)) +
    coord_sf(xlim = c(164.8, 179.4), ylim = c(-47.7, -33.8)) +
    scale_colour_viridis_c() +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("Linear model amplitude residuals (Allen method)")

ggplot() + 
    geom_sf(data = nzmap, fill = "tan1") + 
    geom_sf(data = site.regression.summary.sf,
            mapping = aes(size = phase.pred.err, colour = phase)) +
    coord_sf(xlim = c(164.8, 179.4), ylim = c(-47.7, -33.8)) +
    scale_colour_viridis_c() +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("Linear model phase residuals")

ggplot() + 
    geom_sf(data = nzmap, fill = "tan1") + 
    geom_sf(data = site.regression.summary.sf,
            mapping = aes(size = allen.phase.pred.err, colour = phase)) +
    coord_sf(xlim = c(164.8, 179.4), ylim = c(-47.7, -33.8)) +
    scale_colour_viridis_c() +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("Linear model phase residuals (Allen method)")

ggplot() + 
    geom_sf(data = nzmap, fill = "tan1") + 
    geom_sf(data = site.regression.summary.sf,
            mapping = aes(size = offset.pred.err, colour = offset)) +
    coord_sf(xlim = c(164.8, 179.4), ylim = c(-47.7, -33.8)) +
    scale_colour_viridis_c() +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("Linear model offset residuals")

ggplot() + 
    geom_sf(data = nzmap, fill = "tan1") + 
    geom_sf(data = site.regression.summary.sf,
            mapping = aes(size = allen.offset.pred.err, colour = offset)) +
    coord_sf(xlim = c(164.8, 179.4), ylim = c(-47.7, -33.8)) +
    scale_colour_viridis_c() +
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("Linear model offset residuals (Allen method)")
```




# Linear model of amplitude resisduals from Allen model for regression kriging

Make model, attach it to national grid and plot

```{r}
amp.residuals.mod1 <- lm(allen.amp.pred.err ~  island + ETmp + MSLP + PET + Rad + 
                 RH + SoilM + VP + Wind,
               data = combine.vcsn.reg)

amp.residuals.mod1.step <- MASS::stepAIC(amp.residuals.mod1, direction = "both", trace = FALSE)
summary(amp.residuals.mod1.step)
car::vif(amp.residuals.mod1.step, type = "predictor")
car::Anova(amp.residuals.mod1.step)
national.climate.summary.sf$amp.err.pred<-predict(amp.residuals.mod1, newdata = national.climate.summary)
plot(national.climate.summary.sf["amp.err.pred"], cex = 0.5, pch = 16)
```



# Saved data

## Site summary for estimated sinusoidal parameters

```{r}
save(combine.data, file = "Output/Data/combine.data.RData")
dim(combine.data)     
names(combine.data)
```

## Site summary for estimated sinusoidal parameters and predicted values from regressions

```{r}
save(combine.vcsn.reg, file = "Output/Data/combine.vcsn.reg.RData")
dim(combine.vcsn.reg)     
names(combine.vcsn.reg)

save(national.climate.summary, file = "Output/Data/national.climate.summary.01.output.RData")
dim(national.climate.summary)     
names(national.climate.summary)
```

## RMSE by site

```{r}
save(RMSE.by.site, file = "Output/Data/RMSE.by.site.RData")
dim(RMSE.by.site)     
names(RMSE.by.site)
```
