---
title: "Precipitation sinusoidal isocape model: kriging over NZ of sinusoidal parameters"
author: "Andy McKenzie (NIWA), Bruce Dudley (NIWA)"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output:
  html_document: 
    theme: united
    code_folding: show
    toc: yes
    toc_float:
      collapsed: true   
    fig_caption: yes
    number_sections: yes
    df_print: kable
  word_document: 
    toc: yes
  pdf_document: 
    toc: yes
    fig_caption: yes
    number_sections: yes
editor_options:
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# Global options
#
# echo = TRUE. Show the code
# comment = "". Don't put anything in front of results from code
# warning = FALSE. Don't display warning messages
# cache = TRUE. Only process new code when you knit
# dev = 'png'. Save graphics in figure folder as png
# fig.height = 5. Height of figures (in inches)

unlink("Output/KrigingResults", recursive = TRUE)
dir.create("Output/KrigingResults")

knitr::opts_chunk$set(
  echo = TRUE,
  comment = "",
  warning = FALSE,
  message = FALSE,
  cache = FALSE,
  dev = "png", dpi = 300,
  fig.height = 5
)


# tidyverse packages
library(tidyr)
library(dplyr)
library(readr)
library(ggplot2)
library(lubridate)


# https://docs.stadiamaps.com/guides/migrating-from-stamen-map-tiles/#ggmap
# devtools::install_github("stadiamaps/ggmap”)
library(ggmap) # get_stamenmap() & ggmap
# register_stadiamaps("ff428357-71cc-4a08-8912-9111113265ec", write = TRUE)


# For map plotting
library(sf)
library(stars)

# devtools::install_github("ropensci/rnaturalearthhires")
library(rnaturalearthhires)
library(rnaturalearth)
library(rnaturalearthdata)

library(ggrepel) # geom_text_repel()
library(cowplot) # multiple plots for a figure

# model kriging
library(gstat)

# For reading in global isotope prediction maps
# library(raster)

library(flextable)
set_flextable_defaults(big.mark = "")




rm(list = ls())

source("auxiliary.code.R")

ft <- function(input.data, caption = "", num.rows = Inf) {
  ft <- input.data |>
    slice_head(n = num.rows) |>
    flextable() |>
    autofit() |>
    theme_zebra() |>
    set_caption(caption)
  ft
}
```


# Load in data

```{r}
# site summary for estimated sinusoidal parameters
load("Output/Data/combine.data.RData")
dim(combine.data)  
names(combine.data)

# sites summary for estimated sinusoidal parameters plus predicted
# values from regressions
load("Output/Data/combine.vcsn.reg.RData")
dim(combine.vcsn.reg)  
names(combine.vcsn.reg)

# agent locations (lat/lon)
load("Output/Data/vcsn.agent.locations.RData")
dim(vcsn.agent.locations)  
names(vcsn.agent.locations)

# mean value of climate parameters at agent locations
load("Output/Data/vcsn.mean.RData")
glimpse(vcsn.mean)

# Mean value of climate data at an agent location.
#  - first generated in 00-isoscape-data-preparation.Rmd
#  - modified in 01-isoscape-model-fits-regression.Rmd and output from there
load(file = "Output/Data/national.climate.summary.01.output.RData")
glimpse(national.climate.summary)
```


# Prediction over New Zealand 2 - regression kriging


## Base NZ map

A starting base map of class *sf* is used in a geographic CRS i.e.
coordinates of lat/lon in degrees (not meters)

```{r data base NZ map sf}
nzmap <- rnaturalearth::ne_countries(scale = 'large', 
                      country = "New Zealand",
                      returnclass = 'sf')

st_is_longlat(nzmap)
st_crs(nzmap)$proj4string
```



Crop the starting base map to exclude the Chatham Islands, as above.
Also for Kriging prediction, a projected CRS is needed, and New Zealand
Transverse Mercator 2000 (NZTM 2000) is chosen for this.

```{r data base NZ map in projected CRS}
crs_projected_target <- "EPSG:2193"

nzmap.proj <- nzmap |> 
  st_crop(xmin = 164.8, xmax = 179.4, ymin = -47.7, ymax = -33.8) |> 
  st_transform(crs = crs_projected_target)
```

```{r plot NZ map cropped and projected CRS}
ggplot() +  
  geom_sf(data = nzmap.proj, fill = "darkgreen") +
  ggtitle("Cropped NZ map: NZ Transverse Mercator 2000")
```

Grid for interpolating on. Starting with the projected NZ map, setup a
grip with 5 km grid cells ("pixels")

```{r data nz grid}
nzgrid <- nzmap.proj |> 
  sf::st_bbox() |>
  stars::st_as_stars(dx = 5000) |>
  sf::st_crop(nzmap.proj) 

nzgrid
```



## VCSN data in the projected CRS

The VCSN data set is made into an sf object with the target
CRS: NZTM 2000 (i.e. same as the base NZ map).


```{r data sf format}
combine.vcsn.reg.proj <- st_as_sf(combine.vcsn.reg,
                                    coords = c("VCSN.lon", "VCSN.lat"),
                                    remove = FALSE,
                                    crs = 4326) |> 
                      st_transform(crs = crs_projected_target)  
```


| Name                  | Comment                                                                                   |
|-----------------------|-------------------------------------------------------------------------------------------|
| combine.vcsn          | site, estimated sinusoidal parameters, mean value of VCSN variable over 2007-2009 (inclusive)                                                   |
| combine.vcsn.reg      | removed some VCSN variables that are highly linearly correlated with other VCSN variables |
| combine.vcsn.reg.proj | Turn combine.vcsn.reg into an sf object                                                   |

: What the various VCSN data set are.



## Amplitude prediction

```{r plot NZ project map with amplitude, fig.cap="Estimated amplitude at sites."}
plot.amplitude.spatial <- iso_plot_spatial_static(combine.data, "Site", "lat", "lon", "Amplitude")
plot.amplitude.spatial
```

### Inverse distance weighting

Warm-up with predicted values based on Inverse Distance Weighting (IDW).
Values at sites match the measurements, with inverse weighting for
interpolation elsewhere.

```{r amplitude IDW interpolation}
amp.IDW <- gstat::idw(Amplitude ~ 1, combine.vcsn.reg.proj, nzgrid)
```

```{r plot amplitude IDW}
plot.amplitude.IDW <- 
ggplot() + 
  geom_stars(data = amp.IDW, aes(fill = var1.pred, x = x, y = y)) + 
  geom_sf(data = combine.vcsn.reg.proj) +
  coord_sf(expand = FALSE) +
  scale_fill_continuous(type = "viridis") +
  labs(fill = "Amplitude") +  
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Amplitude: IDW interpolation")  
plot.amplitude.IDW
```


### Kriging constant model

```{r  constant variogram}
v.constant <- gstat::variogram(Amplitude ~ 1, combine.vcsn.reg.proj)
plot(v.constant, plot.numbers = TRUE)
```


Function variogram chooses default for maximum distance (cutoff: one
third of the length of the bounding box diagonal) and (constant)
interval widths (width: cutoff divided by 15).

These defaults can be changed, e.g. by

```{r variogram variation}
v0 <- gstat::variogram(Amplitude ~ 1, 
                       combine.vcsn.reg.proj, 
                       cutoff = 1E6,
                       width = 30000)
plot(v0, plot.numbers = TRUE)
```

Fitting an exponential variogram for all values of h.

```{r plot constant model variogram, eval=FALSE, include=FALSE}
v.constant.model <- gstat::fit.variogram(v.constant, 
                                 vgm(psill = 1, 
                                     model = "Exp", 
                                     range = 500000,
                                     nugget = 0.05))
plot(v.constant, v.constant.model, plot.numbers = TRUE)
```


```{r plot constant model variogram default}
v.constant.model <- gstat::fit.variogram(v.constant, vgm(model = "Exp"))
plot(v.constant, v.constant.model, plot.numbers = TRUE)
# fitted model used in river site analysis to calculate young water fraction
#   ywf = river site amplitude (fit to measured values)/ "river site catchment weighted precipitation amplitude"
#   Where the precipitation amplitude is predicted at points within a river site
#   catchment, then the weighted average by rainfall taken
save(v.constant.model, file = "Output/KrigingResults/v.constant.model.RData")
```

Apply constant kriging model over NZ. 

```{r constant model kriging}
krig.constant <- gstat::krige(Amplitude ~ 1, 
                           combine.vcsn.reg.proj, 
                           nzgrid, 
                           v.constant.model)
save(krig.constant, file = "Output/KrigingResults/krig.constant.RData")
```



```{r plot amplitude kriging}
plot.amplitude.kriging <- 
ggplot() + 
  geom_stars(data = krig.constant, aes(fill = var1.pred, x = x, y = y)) + 
  geom_sf(data = combine.vcsn.reg.proj) +
  coord_sf(expand = FALSE) +
  scale_fill_continuous(type = "viridis") +
  labs(fill = "Amplitude") +  
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Amplitude: ordinary kriging interpolation") 
plot.amplitude.kriging
```



```{r plot amplitude kriging variance}
plot.amplitude.kriging.variance <- 
ggplot() + 
  geom_stars(data = krig.constant, aes(fill = var1.var, x = x, y = y)) + 
  geom_sf(data = combine.vcsn.reg.proj) +
  coord_sf(expand = FALSE) +
  scale_fill_viridis_c(option = "inferno") +
  labs(fill = "Variance") +  
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Amplitude: kriging interpolation variance") 
plot.amplitude.kriging.variance
```




#### Summary plots for constant kriging

```{r plot amplitude IDW and kriging, fig.height=6}
cowplot::plot_grid(plot.amplitude.spatial,  plot.amplitude.IDW,
plot.amplitude.spatial, plot.amplitude.kriging)
```


```{r plot amplitude IDW and kriging variance, fig.height=6}
cowplot::plot_grid(plot.amplitude.kriging, plot.amplitude.kriging.variance)
```




### Kriging trend model

As above, but put in some trend variables (say soil moisture: SoilM) for the
kriging, which needs to assigned to each grid point to predict over NZ. 

This (below) is now done above, no need to repeat here

```{r data read all VCSN data 2007 to 2022}
# It takes a while to retreive the data, so is done once and the data saved
#
# vcsn.2007.2009 <- vcsn_combine_daily_data(VCSN.directory, years = 2007:2009)
# save(vcsn.2007.2009, file = "ProcessedData/vcsn.2007.2009.RData")

#load("output/vcsn.2007.2009.RData")

#vcsn.mean <- vcsn.2007.2009 |>
#  dplyr::select(-Tmin, -TMax, 
#                -Tmin_N, - Tmax_N, 
#                -Rain)  |>
#  rename(lat = Lat, lon = Longt) |>
#  group_by(Agent, lat, lon) |>
#  summarise(across(.cols = MSLP:Rain_bc, .fns = ~mean(.x, na.rm = TRUE))) 

# it's big, best to remove it
#rm(vcsn.2007.2009)

target.crs <- sf::st_crs(nzgrid)

vcsn.mean.sf <- vcsn.mean |>
  sf::st_as_sf(coords = c("lon", "lat"), 
             remove = FALSE,
             crs = 4326) |>
  sf::st_transform(crs = target.crs)
```



```{r data nz find nearest VCSN mean value}
nzgrid.sf <- st_as_sf(nzgrid)
nzgrid.long.format <- st_join(nzgrid.sf, vcsn.mean.sf, join = st_nearest_feature) 

nzgrid.long.format |> slice_head(n = 5)
```


```{r}
v.trend <- gstat::variogram(Amplitude ~ SoilM, combine.vcsn.reg.proj)
plot(v.trend, plot.numbers = TRUE)
```

```{r}
v.trend.model <- gstat::fit.variogram(v.trend, vgm(model = "Exp"))
plot(v.trend, v.trend.model, plot.numbers = TRUE)
save(v.trend.model, file = "Output/KrigingResults/v.trend.model.RData")
```



Apply trend kriging model over NZ. 

```{r trend model kriging}
krig.trend <- gstat::krige(Amplitude ~ SoilM, 
                           combine.vcsn.reg.proj, 
                           nzgrid.long.format, 
                           v.trend.model)
save(krig.trend, file = "Output/KrigingResults/krig.trend.RData")
```


Second universal kriging prediction using Rad and AnnualTempRange

```{r}
v.trend2 <- gstat::variogram(Amplitude ~ Rad + AnnualTempRange, combine.vcsn.reg.proj)
plot(v.trend2, plot.numbers = TRUE)
```

```{r}
v.trend.model2 <- gstat::fit.variogram(v.trend2, vgm(model = "Exp"))
plot(v.trend2, v.trend.model2, plot.numbers = TRUE)
```

```{r trend model kriging2}
amp.krig.trend <- gstat::krige(Amplitude ~ Rad + AnnualTempRange, 
                           combine.vcsn.reg.proj, 
                           nzgrid.long.format, 
                           v.trend.model2)
save(amp.krig.trend, file = "Output/KrigingResults/amp.krig.trend.RData")
```


First universal kriging plot (using SoilM).

```{r plot amplitude kriging with trend}
krig.trend.stars <- stars::st_as_stars(krig.trend)

plot.amplitude.kriging.trend <- 
  ggplot() + 
  geom_stars(data = krig.trend.stars, aes(fill = var1.pred, x = x, y = y)) + 
  geom_sf(data = combine.vcsn.reg.proj) +
  coord_sf(expand = FALSE) +
  scale_fill_continuous(type = "viridis") +
  labs(fill = "Amplitude") +  
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Amplitude: universal kriging interpolation using SoilM") 

plot.amplitude.kriging.trend
```

Second universal kriging plot (using Rad and AnnualTempRange)

```{r plot amplitude kriging with trend2}
krig.trend.stars2 <- stars::st_as_stars(amp.krig.trend)

plot.amplitude.kriging.trend2 <- 
  ggplot() + 
  geom_stars(data = krig.trend.stars2, aes(fill = var1.pred, x = x, y = y)) + 
  geom_sf(data = combine.vcsn.reg.proj) +
  coord_sf(expand = FALSE) +
  scale_fill_continuous(type = "viridis") +
  labs(fill = "Amplitude") +  
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Amplitude: universal kriging interpolation using Rad and TempRange") 

plot.amplitude.kriging.trend2
```

```{r plot compare two universal models for amplitude }
cowplot::plot_grid(plot.amplitude.kriging.trend, plot.amplitude.kriging.trend2)
```


```{r plot compare ordinary vs universal for amplitude II}
cowplot::plot_grid(plot.amplitude.kriging, plot.amplitude.kriging.trend)
```

### Check of predictions for amplitude from universal kriging using SoilM against fitted site amplitudes

```{r}
slice_head(vcsn.agent.locations, n = 3)
```

```{r}
load("Output/KrigingResults/krig.trend.RData")
slice_head(krig.trend, n = 3)
```

Make the kriging simple features object have the same CRS as that of the vcsn agents simple features object

```{r}
agent.CRS <- sf::st_crs(vcsn.agent.locations)

krig.trend.sf <- krig.trend |>
  sf::st_transform(crs = agent.CRS)
```

For the vcsn agent locations find the nearest point on the grid used for the kriging interpolation. 

```{r}
vcsn.agent.locations2 <- vcsn.agent.locations
kriging.row.index <- sf::st_nearest_feature(vcsn.agent.locations, krig.trend.sf)
vcsn.agent.locations2$Amplitude <- krig.trend.sf$var1.pred[kriging.row.index]
slice_head(vcsn.agent.locations2, n = 3)
```
    
 
Check that this looks to be working with some plots. Firstly, the kriging interpolation plot. 

First the amplitude at vcsn agent points. 
    
```{r}
plot(vcsn.agent.locations2["Amplitude"])
```

Now compare the regression modeled amplitude grids against those using regression kriging (using Soil moisture as a predictor). 
    
```{r}
Andyamps <- as.data.frame(vcsn.agent.locations2)
Andyamps <- subset(Andyamps, select = -c(2, 3))
names(Andyamps) <- c("VCSN.Agent", "geometry", "Amplitude.krig")
Andyamps$VCSN.Agent <- as.character(Andyamps$VCSN.Agent)

national.climate.summary$VCSN.Agent <- as.character(national.climate.summary$VCSN.Agent)
combine.vcsn.amps <- dplyr::inner_join(national.climate.summary, Andyamps, by = "VCSN.Agent")
head(combine.vcsn.amps)

ggplot(combine.vcsn.amps, aes(x = amp.pred, y = Amplitude.krig)) +
  geom_point(colour = "blue")
```



## Offset krigging prediction



```{r plot NZ project map with offset, fig.cap="Estimated offset at sites."}
plot.offset.spatial <- iso_plot_spatial_static(combine.data, "Site", "lat", "lon", "offset")
plot.offset.spatial
```


```{r  constant variogram offset}
v.constant.offset <- gstat::variogram(offset ~ 1, combine.vcsn.reg.proj)
plot(v.constant, plot.numbers = TRUE)

v.constant.model.offset <- gstat::fit.variogram(v.constant.offset, vgm(model = "Exp"))
plot(v.constant.offset, v.constant.model.offset, plot.numbers = TRUE)
```

Apply constant kriging model over NZ. 

```{r}
krig.constant.offset <- gstat::krige(offset ~ 1, 
                           combine.vcsn.reg.proj, 
                           nzgrid, 
                           v.constant.model.offset)
```



```{r plot offset kriging}
plot.offset.kriging <- 
ggplot() + 
  geom_stars(data = krig.constant.offset, aes(fill = var1.pred, x = x, y = y)) + 
  geom_sf(data = combine.vcsn.reg.proj) +
  coord_sf(expand = FALSE) +
  scale_fill_continuous(type = "viridis") +
  labs(fill = "Offset") +  
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Offset: ordinary kriging interpolation") 
plot.offset.kriging
```


```{r plot offset kriging variance}
plot.offset.kriging.variance <- 
ggplot() + 
  geom_stars(data = krig.constant.offset, aes(fill = var1.var, x = x, y = y)) + 
  geom_sf(data = combine.vcsn.reg.proj) +
  coord_sf(expand = FALSE) +
  scale_fill_viridis_c(option = "inferno") +
  labs(fill = "Variance") +  
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Offset: kriging interpolation variance") 
plot.offset.kriging.variance
```



Then universal kriging prediction using best VCSN predictors from tree regressions (VP, AnnualTempRange, ETmp and SoilM)


```{r}
v.trend.offset <- gstat::variogram(offset ~ VP + AnnualTempRange + ETmp + SoilM,
                             combine.vcsn.reg.proj)
v.trend.offset.model <- gstat::fit.variogram(v.trend.offset, 
                                             vgm(model = "Exp"))
plot(v.trend.offset, v.trend.offset.model, plot.numbers = TRUE)
```


```{r}
offset.krig.trend <- gstat::krige(offset ~ VP + AnnualTempRange + ETmp + SoilM, 
                           combine.vcsn.reg.proj, 
                           nzgrid.long.format, 
                           v.trend.offset.model)
save(offset.krig.trend, file = "Output/KrigingResults/offset.krig.trend.RData")
```


Universal kriging plot (using VP + AnnualTempRange + ETmp + SoilM)

```{r}
offset.krig.trend.stars <- stars::st_as_stars(offset.krig.trend)

plot.offset.kriging.trend <- 
  ggplot() + 
    geom_stars(data = offset.krig.trend.stars, aes(fill = var1.pred)) + 
    geom_sf(data = combine.vcsn.reg.proj) +
    coord_sf(expand = FALSE) +
    scale_fill_continuous(type = "viridis") +
    labs(fill = "Offset") +  
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("Offset: universal kriging interpolation")

plot.offset.kriging.trend
```


```{r plot offset fitted and kriging, fig.height=6}
cowplot::plot_grid(plot.offset.spatial,  plot.offset.kriging, plot.offset.kriging.trend)
```





## Phase krigging prediction



```{r plot NZ project map with phase, fig.cap="Estimated phase at sites."}
plot.phase.spatial <- iso_plot_spatial_static(combine.data, "Site", "lat", "lon", "phase")
plot.phase.spatial
```



```{r  constant variogram phase}
v.constant.phase <- gstat::variogram(phase ~ 1, combine.vcsn.reg.proj)
plot(v.constant, plot.numbers = TRUE)

v.constant.model.phase <- gstat::fit.variogram(v.constant.phase, vgm(model = "Exp"))
plot(v.constant.phase, v.constant.model.phase, plot.numbers = TRUE)
```

Apply constant kriging model over NZ. 

```{r constant model kriging phase}
krig.constant.phase <- gstat::krige(phase ~ 1, 
                           combine.vcsn.reg.proj, 
                           nzgrid, 
                           v.constant.model.phase)
```



```{r plot phase kriging}
plot.phase.kriging <- 
ggplot() + 
  geom_stars(data = krig.constant.phase, aes(fill = var1.pred, x = x, y = y)) + 
  geom_sf(data = combine.vcsn.reg.proj) +
  coord_sf(expand = FALSE) +
  scale_fill_continuous(type = "viridis") +
  labs(fill = "phase") +  
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Phase: ordinary kriging interpolation") 
plot.phase.kriging
```


```{r plot phase kriging variance}
plot.phase.kriging.variance <- 
ggplot() + 
  geom_stars(data = krig.constant.phase, aes(fill = var1.var, x = x, y = y)) + 
  geom_sf(data = combine.vcsn.reg.proj) +
  coord_sf(expand = FALSE) +
  scale_fill_viridis_c(option = "inferno") +
  labs(fill = "Variance") +  
  xlab("Longitude") +
  ylab("Latitude") +
  ggtitle("Phase: kriging interpolation variance") 
plot.offset.kriging.variance
```


Then universal kriging prediction using best VCSN predictors from tree regressions (VP and height)



```{r}
v.trend.phase <- gstat::variogram(phase ~ VP + height,
                             combine.vcsn.reg.proj)
v.trend.phase.model <- gstat::fit.variogram(v.trend.phase,  
                                             vgm(model = "Exp"))
plot(v.trend.phase, v.trend.phase.model, plot.numbers = TRUE)
```



```{r trend model kriging for offset}
phase.krig.trend <- gstat::krige(phase ~ VP + height, 
                           combine.vcsn.reg.proj, 
                           nzgrid.long.format, 
                           v.trend.phase.model)
save(phase.krig.trend, file = "Output/KrigingResults/phase.krig.trend.RData")
```


A variation on the trend variables used. 

```{r}
v.trend2 <- gstat::variogram(phase ~ VCSN.lat + VCSN.lon, combine.vcsn.reg.proj)
plot(v.trend2, plot.numbers = TRUE)

v.trend2.phase <- gstat::fit.variogram(v.trend2, vgm(model = "Exp"))
plot(v.trend2, v.trend2.phase, plot.numbers = TRUE)
```


Universal kriging plot (phase) using VP and height.

```{r plot offset kriging with trend}
phase.krig.trend.stars <- stars::st_as_stars(phase.krig.trend)

plot.phase.kriging.trend <- 
  ggplot() + 
    geom_stars(data = phase.krig.trend.stars, aes(fill = var1.pred)) + 
    geom_sf(data = combine.vcsn.reg.proj) +
    coord_sf(expand = FALSE) +
    scale_fill_continuous(type = "viridis") +
    labs(fill = "Phase") +  
    xlab("Longitude") +
    ylab("Latitude") +
    ggtitle("Phase: universal kriging interpolation") 

plot.phase.kriging.trend
```

Some phase maps made using the two different kriging  approaches

```{r plot phase fittted and kriging, fig.height=6}
cowplot::plot_grid(plot.phase.spatial, 
                   plot.phase.kriging,
                   plot.phase.kriging.trend)
```

#Plots of all three regression kriging interpolations: Amplitude, phase and offset. 
```{r}
cowplot::plot_grid(plot.amplitude.kriging.trend, 
                   plot.phase.kriging.trend,
                   plot.offset.kriging.trend)
```




#Add kriging model predictions to VCSN grid

First make .sf objects

```{r}
agent.CRS <- sf::st_crs(vcsn.agent.locations)

amp.krig.trend.sf <- krig.trend |>
  sf::st_transform(crs = agent.CRS)

phase.trend.sf <- phase.krig.trend |>
  sf::st_transform(crs = agent.CRS)

offset.trend.sf <- offset.krig.trend |>
  sf::st_transform(crs = agent.CRS)
```

Then the extraction

```{r}
vcsn.agent.locations3<-vcsn.agent.locations
kriging.row.index1 <- sf::st_nearest_feature(vcsn.agent.locations, amp.krig.trend.sf)
vcsn.agent.locations3$Amplitude <- amp.krig.trend.sf$var1.pred[kriging.row.index1]

kriging.row.index2 <- sf::st_nearest_feature(vcsn.agent.locations, phase.trend.sf)
vcsn.agent.locations3$Phase <- phase.trend.sf$var1.pred[kriging.row.index2]

kriging.row.index3 <- sf::st_nearest_feature(vcsn.agent.locations, offset.trend.sf)
vcsn.agent.locations3$Offset <- offset.trend.sf$var1.pred[kriging.row.index3]

slice_head(vcsn.agent.locations3, n = 3)
```



Then attach kriging predictions to the data frame of climate data and predictions from linear models 
    
```{r}
national.climate.summary$krig.amp.pred    <- vcsn.agent.locations3$Amplitude
national.climate.summary$krig.phase.pred  <- vcsn.agent.locations3$Phase
national.climate.summary$krig.offset.pred <- vcsn.agent.locations3$Offset

head(national.climate.summary)
plot(national.climate.summary$krig.amp.pred, national.climate.summary$amp.pred)
```


# Saved data


```{r}
save(national.climate.summary, file = "Output/Data/national.climate.summary.02.output.RData")
dim(national.climate.summary)
names(national.climate.summary)
```


```{r}
save(nzgrid.long.format, file = "Output/Data/nzgrid.long.format.RData")
glimpse(nzgrid.long.format)
```

